Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach

### 2013-10-09 inst/doc/nonparametric-bayes.Rmd
compile draft
writing

<!-- Uncertainty outside the data without the correct model has not been handled. -->

@@

<!-- What do we call this?  Extrapolation uncertainty?  Pathological Uncertainty? -->

@@

@@

<!-- More Consise versions above

- Parametric uncertainty

@@

Resource management and conservation planning seek to determine the
optimal set of feasible actions to maximize the value of some objectives
(see [@Halpern2012])  Process error, measurement error, implementation
error [@Reed1979, @Clark1986, @Roughgarden1996, @Sethi2005].  These
sources of stochasticity in turn mean that model parameters can only
be estimated approximately, requiring parametric uncertainty also be
considered [@Ludwig1982, ].

## Two central challenges:

- We don't know the model
- We don't have data from where we need it most

## Why we don't know the model

- Complex dynamics  [@Glaser2013](http://doi.org/10.1111/faf.12037 "Complex dynamics may limit prediction in marine fisheries")

- model choice and model averaging approaches

## Why we don't have data where we need it

- Concerns of tipping points 
- Danger of learning

- _MAP_

<!-- necessary? -->
This paper represents the first time the SDP decision-making framework has been used
without an a priori model of the underlying dynamics through the use of the BNP approach.

In contrast to parametric models which can only reflect uncertainty 
in parameter estimates, the BNP approach provides a more state-space dependent representation
of uncertainty.

This permits a much greater uncertainty far from the observed data than near the observed 
data.  These features allow the GP-SDP approach to find robust management solutions 
in face of limited data and without knowledge of the correct model structure.

>>>>

<!-- Uncertainty outside the data without the correct model has not been handled. -->
<!-- __We don't have the model__ -->

@@

<!-- __We don't have the data where we need it__ -->
<!-- What do we call this?  Extrapolation uncertainty?  Pathological Uncertainty? -->

@@

<!-- 
## Why we don't know the model
- Complex dynamics  [@Glaser2013](http://doi.org/10.1111/faf.12037 "Complex dynamics may limit prediction in marine fisheries")
- model choice and model averaging approaches
-->

<!--
## Why we don't have data where we need it
- Concerns of tipping points 
- Danger of learning 
-->

@@

<!-- More Consise versions are now above 
- Parametric uncertainty

@@

Resource management and conservation planning seek to determine the
optimal set of feasible actions to maximize the value of some objectives
(see [@Halpern2012])  Process error, measurement error, implementation
error [@Reed1979, @Clark1986, @Roughgarden1996, @Sethi2005].  These
sources of stochasticity in turn mean that model parameters can only
be estimated approximately, requiring parametric uncertainty also be
considered [@Ludwig1982, ].
-->

<!-- necessary? --> <!-- should reflect the dual problem of extrapolation and model uncertainty better -->

This paper represents the first time the SDP decision-making framework has
been used without an a priori model of the underlying dynamics through
the use of the BNP approach.  In contrast to parametric models which
can only reflect uncertainty in parameter estimates, the BNP approach
provides a more state-space dependent representation of uncertainty.
This permits a much greater uncertainty far from the observed data than
near the observed data.  These features allow the GP-SDP approach to
find robust management solutions in face of limited data and without
knowledge of the correct model structure.
